---
title: "get_fish_info"
output: html_notebook
---

Documentation for rfishbase: (https://cran.r-project.org/web/packages/rfishbase/vignettes/tutorial.html)

# Load libraries

Load the libraries we need.

```{r}
library(rfishbase)
library(tidyverse)
library(ggplot2)
library(Hmisc)
```

# Test a few species

NB: Don't call the list of species `species` because that's the function we need to call in rfishbase to get information.
```{r}
fish <- c("Anoplogaster cornuta",
             "Platichthys stellatus")
```

```{r}
validate_names(fish)
```

```{r}
fishdata <- species(fish)
```

```{r}
str(fishdata)
```

```{r}
dat <- species(fish, fields=c("BodyShapeI", "DepthRangeShallow", "DepthRangeDeep", "DepthRangeComShallow", "DepthRangeComDeep", "Length", "CommonLength", "Weight", "DemersPelag", "Subfamily"))
```

```{r}
dat
```

# Load species from CSV file
```{r}
filename <- "../Merged Data/VertMeasurements_Shape-Habitat.csv"
```

```{r}
vertmeas <- read.csv(filename, colClasses = c("Species"="character"))

vertmeas
```

```{r}
str(vertmeas)
```

```{r}
#Idea for removing underscores, looked at https://swcarpentry.github.io/r-novice-inflammation/11-supp-read-write-csv/index.html

vertmeas <- mutate(vertmeas,
                   Species = sub("_", " ", Species)) #ifelse(as.character(Species)=='Alectis_ciliaris', 'Dana', as.character(Species)))
# vertmeas$Species <- as.factor(vertmeas$Species)
```

```{r}
vertmeas$Species
```

```{r}
fish <- vertmeas %>% distinct(Species)
```

```{r}
fish$Species
```

#Create Data Frame for Fishbase data for Measured Species 
New challenge: Find the missing species and correct them
Missing: Roeboides affinis (on Fishbase), Stenobrachius leucopsaurus (on Fishbase), Onchorhyncus gorbuscha (not working), Ammodytes personatus (on Fishbase), Anoplarchus pupurescens (misspelled)

--> maybe because not all indicated data points we want are there for these species

```{r}
validate_names(fish$Species)
```

```{r}
fishbasedata <- species(fish$Species, fields=c("BodyShapeI", "DepthRangeShallow", "DepthRangeDeep", "DepthRangeComShallow", "DepthRangeComDeep", "Length", "CommonLength", "Weight", "DemersPelag", "Subfamily"))
```

```{r}
fishbasedata
```


```{r}
str(fishbasedata)
```

#Merge Depth Columns of Fishbase Data Into New Columns
```{r}
fishbasedatamergeddepth <- fishbasedata %>%
  mutate(newdepthrangeshallow = replace(DepthRangeComShallow,
                           is.na(DepthRangeComShallow),DepthRangeShallow),
         newdepthrangedeep = replace(DepthRangeComDeep,
                          is.na(DepthRangeComDeep),DepthRangeDeep))
fishbasedatamergeddepth
```

```{r}
write.csv(fishbasedatamergeddepth, "fishbasedata_mergeddepth.csv")
```


```{r}
str(fishbasedatamergeddepth)
```

#Create csv with the Fishbase Data for Each Species
```{r}
getwd()
```


```{r}
write.csv(fishbasedata,"fishbasedata.csv")

#couldn't get it to name the rows each species name so might have to go through and put that manually but I now have that csv
#called fishbasedata.csv
#manually named each column in Excel sheet called fishbasedata_measuredspecies.xlsx
```

```{r}
help(rfishbase)
```

#Attempt to see if able to find species that weren't validated 
```{r}
missingfish <- c("Roeboides affinis","Stenobrachius leucopsaurus","Ammodytes personatus")
```

```{r}
validate_names(missingfish)
```

```{r}
missingfishdata <- species(missingfish)
```

```{r}
str(missingfishdata)
```

```{r}
missingdat <- species(missingfish, fields=c("BodyShapeI", "DepthRangeShallow", "DepthRangeDeep", "DepthRangeComShallow", "DepthRangeComDeep", "Length", "CommonLength", "Weight", "DemersPelag", "Subfamily"))
```

```{r}
missingdat
```

#Merge Fishbase Data and Vertebrae Measurements 
```{r}
fishbasedata2 <- cbind(fish, fishbasedatamergeddepth)
```

```{r}
left_join(
  vertmeas %>% select(Species, Indiv, SL, Pos, CBL, d, alpha_Pos, D_Pos, alpha_Ant, D_Ant) %>%
  pivot_wider(names_from = Pos, values_from = c("CBL", "d", "alpha_Pos", "D_Pos", "alpha_Ant", "D_Ant")),
  
  fishbasedata2,
  
  by = "Species")
```

```{r}
vertmeasmean <- vertmeas %>%
  group_by(Species, Indiv) %>%
  select(SL, CBL, d, alpha_Pos, alpha_Ant, D_Pos, D_Ant) %>%
  summarize_all(mean, na.rm=TRUE) %>%
  ungroup()

vertmeasmean
```

```{r}
vertmeasmeans_fishbase <-
  left_join(
  vertmeasmean,
  
  fishbasedata2,
  
  by = "Species")
```

```{r}
vertmeasall_fishbase <-
  left_join(
  vertmeas,
  
  fishbasedata2,
  
  by = "Species")

vertmeasall_fishbase
```


#Exploratory Scatter Plots for vertmeasmeans_fishbase
TO DO: Write text for exploratory plots and save out vertmeasmeans_fishbase

```{r}
write.csv(vertmeasmeans_fishbase, "vertmeasmeans_fishbase.csv")
```


```{r}
ggplot(vertmeasmeans_fishbase, aes(x = CBL, y = d, color = DemersPelag)) +
       geom_point()
  # facet_wrap(~ DemersPelag)
```
Here there seems to be a positive relationship with only some subsets of the DemersPelag category with these species. Mainly the demersal seem to have a positive relationship between the CBL and d, as well as some of the benthopelagic species. 


```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = D_Pos, color = Length)) +
       geom_point() 
#facet_wrap(~Length)
```
Here there does not seem to be as much of a trend, because the lengths of these fish are more variable than a category like DemersPelag or depth range. However, some of the shorter fish species do seem to have a positive relationships with d and D_Pos, but this trend is hard to tell for sure. 


```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = alpha_Pos, color = newdepthrangedeep)) + geom_point() 
#facet_wrap(~newdepthrangedeep)
```
In this plot, generally d and alpha_Pos seem to have a positive relationship when categorized by the deepest limits of their depth range (this column being the merged common and full depth range value columns). This trend seems to be more readily seen with the fish for which the depth range data were unavailable. 


```{r}
ggplot(vertmeasmeans_fishbase, aes(x = CBL, y = D_Ant, color = BodyShapeI)) + geom_point()
```
Unsure of why it removed this many rows... Still displays a positive relationships between CBL and D_Ant for all body shape types present. This trend is most easily seen with the "short and/or deep" body shape.


#Attempt to figure out why plots with D_Ant were causing issues 
```{r}
vertmeasmeans_fishbase %>%
  group_by(Species) %>%
  filter(is.na(D_Ant))
```
D_Ant and alpha_Ant don't appear for a large chunk of these species because with the elongate fish that were Cassandra's measurements, she only measured D_Pos and alpha_Pos, not the anterior ones. 

#More Exploratory Scatter Plots 
```{r}
ggplot(vertmeasall_fishbase, aes(x = CBL, y = D_Ant, color = BodyShapeI)) + geom_point()
```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = D_Pos, color = DemersPelag)) + geom_point()
```
Here, there also seems so be somewhat of a positive relationship between d and D_Pos when colored by DemersPelag. This positive relationship looks steepr with the reef-associated category than with the demersal group, but both seem to exhibit somewhat of a positive relationship.

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = alpha_Pos, color = Weight)) + geom_point()
```
Here there is a positive relationship between d and alpha_Pos, but there are many N/A in the weight data, but the lighter fish still had a positive relationship with alpha_Pos and d. 

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = CBL, color = BodyShapeI)) + geom_point() 
#facet_wrap(~BodyShapeI)
```
Here there is less of a positive relationship between the CBL and d for some of these species. The only category that seems to display this positive relationship is the eel-like body shape. CBL

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = alpha_Pos, y = D_Pos, color = BodyShapeI)) + geom_point() 
#facet_wrap(~BodyShapeI)
```
Here there seems to be a positive relationship displayed mainly with the eel-like and elongated fishes. There are not as many N/A values here as there are in other categories. 

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = alpha_Pos, y = D_Pos, color = DemersPelag)) + geom_point() 
#facet_wrap(~DemersPelag)
```
With the same x and y but colored differently (this time by DemersPelag), the positive relationship is displayed in mainly the demersal, benthopelagic, and reef-associated fishes. The reef-associated fishes have mostly larger D_Pos and smaller alpha_Pos, with the demersal fishes having the opposite. 

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = d, y = D_Ant, color = newdepthrangedeep)) + geom_point() 
#facet_wrap(~newdepthrangedeep)
```
Many missing values in the depth columns (even in the merged ones). Here this is displaying mainly fish species whose lowest part of the depth range is not as shallow, but there are more deep sea fishes present, shown in the lighter blue colors. 

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = CBL, y = alpha_Pos, color = Length)) + geom_point()
```
There does not seem to be much relationship between CBL and alpha_Pos, but all of the shorter species seem to have smaller alpha_Pos, which would make sense due to their smaller size.

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = alpha_Ant, y = D_Ant, color = DemersPelag)) + geom_point()
```
This for some reason has many rows taken out? I'm not sure why... but within each DemersPelag category there seems to be somewhat of a positive relationship between D_Ant and alpha_Ant.


#Mean plots with error bars for more exploratory plots 
```{r}
ggplot(vertmeasmeans_fishbase, aes(x = DemersPelag, y = d, color=DemersPelag)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position="jitter", alpha=0.5)
```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = BodyShapeI, y = d, color=BodyShapeI)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position="jitter", alpha=0.5)
```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = DemersPelag, y = D_Pos, color=DemersPelag)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position ="jitter", alpha=0.5)
```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = DemersPelag, y = alpha_Ant, color = DemersPelag))+
  stat_summary(fun.data = "mean_cl_boot")+
  geom_point(position = "jitter", alpha=0.5)

```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = BodyShapeI, y = alpha_Ant, color = BodyShapeI)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position = "jitter", alpha=0.5)
```

```{r}
ggplot(vertmeasmeans_fishbase, aes(x = BodyShapeI, y = CBL, color = BodyShapeI)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position = "jitter", alpha=0.5)
```

