---
title: "Main"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(plotly)
library(knitr)
library(here)
```

```{r}
here(i_am('Code/Main.Rmd'))
```

```{r}
options(knitr.duplicate.label = "allow")
```

# Raw data

The initial measurements were made by Dana Baxter by clicking points on vertebrae from CT scans. The measurements are in `MasterVert_Measurements_old.csv`.

Dana processed the data from Cymatogaster aggregata a bit later, so we have to merge it into the main data set.

Input: `Raw Data/MasterVert_Measurements_old.csv`
Output: `Raw Data/MasterVert_Measurements.csv`
```{r}
rmarkdown::render(here('Code/ProcessCymatogasterFiles.Rmd'))
```

# Matching species

Species are distributed somewhat randomly throughout the bony fishes. We use the phylogeny from Betancur-R et al 2017. Not every species we digitized is in their tree. For most of them, we can use another species in the same genus, since we don't have many species in our data set that are in one genus.

Input: `Raw Data/MasterVert_Measurements.csv`
Output: `Processed Data/MasterVert_Measurements_Matched.csv`
   This contains all of the measurements for all of the vertebrae, but with a new set of columns for "MatchSpecies" and "MatchGenus", which are the matching species in the Betancur-R tree.
```{r}
rmarkdown::render(here('Code/CheckSpeciesMatch.Rmd'))
```

# Summary data sets

## Centering and taking maxes and medians

Here, we take the data for each vertebra and produce summary values for each species. We first center the data set by finding the median value of any measurement for each vertebra across all of the species, then subtracting the actual value from that median.

Input: `Processed Data/MasterVert_Measurements_Matched.csv`
Output: 

* `Processed Data/vertdata_centered.csv`: This contains the centered measurements
* `Processed Data/vertdata_summary.csv`: This contains the summary maximum values and medians for each species
```{r}
rmarkdown::render(here('Code/summarize_vert_meas.Rmd'))
```


## Fitting the data with linear models

To represent the variation of the measurements along the vertebrae, we fit linear models to each measurement as a function of position along the body. Based on AIC, we use just an overall mean, a slope and a midpoint, or a quadratic fit.

We use midpoint rather than intercept. Midpoint is the value at 80% of the way down the body. If we use intercept, we sometimes end up with negative values, which is weird.

Input: `Processed Data/vertdata_centered.csv`,
   `Processed Data/vertdata_summary.csv`
Output: 

* `Processed Data/vertdata_summary_lm.csv`: Contains the midpoint, slope, and quadratic coefficients as appropriate for each model.
* `Processed Data/vertdata_summary_lm_models.Rds`: Same data, but storing the actual models in an Rds file.
```{r}
rmarkdown::render(here('Code/VertLM.Rmd'))
```

# Statistics

Here, we run the phylogenetic generalized least squares models to determine which variables differ among species.

Input: `Processed Data/vertdata_centered.csv`,
   `Processed Data/vertdata_summary.csv`
Output: 

* `Processed Data/vertdata_summary_lm_species.csv`: Contains the summary data with the matched tip and label from the phylogenetic tree
* `Processed Data/vert_tree.rds`: The phylogeny containing just our species in RData format. The CSV export wasn't working properly for tree data
```{r}
rmarkdown::render(here('Code/VertPGLS.Rmd'))
```

# Figures

Figure 1: Shows the phylogeny and several example vertabrae.

This produces the phylogeny.
```{r}
rmarkdown::render(here('Code/plot_phylogeny.Rmd'))
```



