---
title: "PGLSforeachfeature"
output: html_notebook
---

```{r}
library(rfishbase)
library(tidyverse)
library (ggplot2)
library(Hmisc)
library(phytools)
library(here)
library(ape)
library(geiger)
library(nlme)
library(patchwork)
library(emmeans)
library(multcompView)
library(readxl)
library(plotly) 
library(ggthemes) 
library(naniar)
library(dplyr)
library(FactoMineR)
library(factoextra)
library(ggExtra)
library(GGally)
```

#PCA with each individual measurement feature
```{r}
vertdata <- read_csv("All_Vert_Data.csv")
```

```{r}
head(vertdata)
```

```{r}
p1 <- ggplot(vertdata, aes(x=CBL, color=Habitat)) + geom_density()
p2 <- ggplot(vertdata, aes(x=d, color=Habitat)) + geom_density()
p3 <- ggplot(vertdata, aes(x=alpha_Ant, color=Habitat)) + geom_density()
p4 <- ggplot(vertdata, aes(x=alpha_Pos, color=Habitat)) + geom_density()

(p1 | p2) / (p3 | p4)
```

```{r}
vertdata <-
  vertdata %>%
  select(c('Genus','Species', 'Indiv', 'Body.Shape', 'Habitat', 'Tip', 'Family', 'FullName',
           'Pos', 'CBL', 'd', 'alpha_Pos', 'alpha_Ant')) %>%
  # "complete" the Pos column by filling in values that are missing for Species
  # and individuals but present for others
  complete(Pos, nesting(Species, Indiv)) %>%
  arrange(Species, Indiv, Pos) %>%
  rename(alphaPos = alpha_Pos,
         alphaAnt = alpha_Ant)

vertdata
```

```{r}
vertdata <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  fill(Genus, Body.Shape, Habitat, Tip, Family, FullName, .direction='downup') %>%
  ungroup()

vertdata
```

```{r}
vertdata <-
  vertdata %>%
  unite("Species", c("Genus", "Species")) %>%
  relocate(Species, Indiv, Habitat, Tip, Pos, CBL:alphaAnt)

vertdata
```

```{r}
vertdata_full <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  mutate_at(vars(CBL, d, alphaPos, alphaAnt), ~ replace(.x, is.na(.x), mean(.x, na.rm=TRUE))) %>%
  ungroup()

vertdata_mean <- vertdata_full %>%
  group_by(Species, Indiv) %>%
  select(CBL, d, alphaPos, alphaAnt) %>%
  summarize_all(mean, na.rm=TRUE) %>%
  ungroup()


vertdata_full <- left_join(
  vertdata_full,
  vertdata_mean,
  by = "Species", "Indiv")

vertdata_full

#d.y is the parameter with the means
```

Dana, get mean d, CBL, alphaAnt here, before pivot

```{r}
vertdata %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'alphaAnt')) %>%
  summarize_if(is.numeric, ~ sum(!is.na(.x)) / n()) %>%
  select_if(~ .x < 0.8)
```

```{r}
vertdata_wide <-
  vertdata_full %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'alphaAnt'))

vertdata_wide
```

```{r}
vertdata.pca_d <- 
  vertdata_wide %>%
  select(-contains(c("20", "30"))) %>%
  select(-starts_with("CBL")) %>%
  select(-starts_with("alphaPos")) %>%
  select(-starts_with("alphaAnt")) %>%
  PCA(quali.sup=seq(1,7), scale.unit = TRUE)

vertdata.pca_d
```
*From Eric*: If we're working with just d (or just CBL or alphaAnt), then it produces the same results if we have `scale.unit = TRUE` or `scale.unit = FALSE`.

```{r}
vertdata.pca_CBL <- 
  vertdata_wide %>%
  select(-contains(c("20", "30"))) %>%
  select(-starts_with("d")) %>%
  select(-starts_with("alphaPos")) %>%
  select(-starts_with("alphaAnt")) %>%
  PCA(quali.sup=seq(1,7), scale.unit = TRUE)

vertdata.pca_CBL
```

```{r}
vertdata.pca_alphaPos <- 
  vertdata_wide %>%
  select(-contains(c("20", "30"))) %>%
  select(-starts_with("d")) %>%
  select(-starts_with("CBL")) %>%
  select(-starts_with("alphaAnt")) %>%
  PCA(quali.sup=seq(1,7), scale.unit = TRUE)

vertdata.pca_alphaPos
```
What does this mean? Why does it look so drastically different from the others? Is it that Dim1 holds so much more of the variation? But there are others with Dim1 with %s in the 80s.

```{r}
fviz_eig(vertdata.pca_d, addlabels=TRUE)
```

```{r}
fviz_eig(vertdata.pca_CBL, addlabels=TRUE)
```

```{r}
fviz_eig(vertdata.pca_alphaPos, addlabels=TRUE)
```

```{r}
vertdata.var_d <- get_pca_var(vertdata.pca_d)
vertdata.var_d
```

```{r}
vertdata.var_CBL <- get_pca_var(vertdata.pca_CBL)
vertdata.var_CBL
```

```{r}
vertdata.var_alphaPos <- get_pca_var(vertdata.pca_alphaPos)
vertdata.var_alphaPos
```

```{r}
vertdata.var_d$coord
```

```{r}
vertdata.var_CBL$coord
```

```{r}
vertdata.var_alphaPos$coord
```

```{r}
pcweights_d <-
  vertdata.var_d$coord %>%
  as.data.frame() %>%
  rownames_to_column("varpos") %>%
  separate(varpos, into=c("var", "Pos"), sep="_") %>%
  pivot_longer(Dim.1:Dim.5, names_to="PC", values_to="weight") %>%
  pivot_wider(names_from="var", values_from="weight") %>%
  mutate(Pos = as.numeric(Pos),
         PC = as.factor(PC)) %>%
  arrange(PC, Pos)

pcweights_d
```

```{r}
pcweights_CBL <-
  vertdata.var_CBL$coord %>%
  as.data.frame() %>%
  rownames_to_column("varpos") %>%
  separate(varpos, into=c("var", "Pos"), sep="_") %>%
  pivot_longer(Dim.1:Dim.5, names_to="PC", values_to="weight") %>%
  pivot_wider(names_from="var", values_from="weight") %>%
  mutate(Pos = as.numeric(Pos),
         PC = as.factor(PC)) %>%
  arrange(PC, Pos)

pcweights_CBL
```

```{r}
pcweights_alphaPos <-
  vertdata.var_alphaPos$coord %>%
  as.data.frame() %>%
  rownames_to_column("varpos") %>%
  separate(varpos, into=c("var", "Pos"), sep="_") %>%
  pivot_longer(Dim.1:Dim.5, names_to="PC", values_to="weight") %>%
  pivot_wider(names_from="var", values_from="weight") %>%
  mutate(Pos = as.numeric(Pos),
         PC = as.factor(PC)) %>%
  arrange(PC, Pos)

pcweights_alphaPos
```

```{r}
pcweights_d <-
  pcweights_d %>%
  group_by(PC) %>%
  mutate_at(vars(d), ~ .x * sign(mean(.x))) %>%
  ungroup()
```

```{r}
pcweights_CBL <-
  pcweights_CBL %>%
  group_by(PC) %>%
  mutate_at(vars(CBL), ~ .x * sign(mean(.x))) %>%
  ungroup()
```

```{r}
pcweights_alphaPos <-
  pcweights_alphaPos %>%
  group_by(PC) %>%
  mutate_at(vars(alphaPos), ~ .x * sign(mean(.x))) %>%
  ungroup()
```

```{r}
pcweights_d %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=d, color=PC)) +
  geom_line()
```

```{r}
pcweights_CBL %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=CBL, color=PC)) +
  geom_line()
```

```{r}
pcweights_alphaPos %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=alphaPos, color=PC)) +
  geom_line()
```

```{r}
vertdata.ind_d <- get_pca_ind(vertdata.pca_d)

pcscores_d <-
  vertdata.ind_d$coord %>%
  as.data.frame() %>%
  bind_cols(vertdata_wide)

pcscores_d
```

```{r}
vertdata.ind_CBL <- get_pca_ind(vertdata.pca_CBL)

pcscores_CBL <-
  vertdata.ind_CBL$coord %>%
  as.data.frame() %>%
  bind_cols(vertdata_wide)

pcscores_CBL
```

```{r}
vertdata.ind_alphaPos <- get_pca_ind(vertdata.pca_alphaPos)

pcscores_alphaPos <-
  vertdata.ind_alphaPos$coord %>%
  as.data.frame() %>%
  bind_cols(vertdata_wide)

pcscores_alphaPos
```

```{r}
ggpairs(pcscores_d, columns = c("Dim.1", "Dim.2", "Dim.3"),
        mapping = aes(color = Habitat, shape = Habitat),
        upper = list(continuous = "blank"),
        diag = list(continuous = "densityDiag"))
```

```{r}
ggpairs(pcscores_CBL, columns = c("Dim.1", "Dim.2", "Dim.3"),
        mapping = aes(color = Habitat, shape = Habitat),
        upper = list(continuous = "blank"),
        diag = list(continuous = "densityDiag"))
```

```{r}
ggpairs(pcscores_alphaPos, columns = c("Dim.1", "Dim.2", "Dim.3"),
        mapping = aes(color = Habitat, shape = Habitat),
        upper = list(continuous = "blank"),
        diag = list(continuous = "densityDiag"))
```


#PGLS with each individual measurement feature's PCA

```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```

```{r}
tree <- read.tree(here('..', treefile))
```

```{r}
plotTree(tree)
```

```{r}
allspecies <- as.data.frame(tree$tip.label, stringsAsFactors = FALSE) 
colnames(allspecies) <- c('FullName')

allspecies <- 
  allspecies %>% separate(FullName, sep='_', into=c('Family', 'Genus', 'Species'), 
                          extra='drop', remove=FALSE)

allspecies$Tip <- seq_len(nrow(allspecies))
```

```{r}
vertdata
```

```{r}
pcscores_d
```

```{r}
pcscores_CBL
```

```{r}
pcscores_alphaPos
```

```{r}
pcscores_edit_d <- pcscores_d %>% 
  filter(Indiv == "1") 

pcscores_edit_d
```

```{r}
pcscores_edit_CBL <- pcscores_CBL %>% 
  filter(Indiv == "1") 

pcscores_edit_CBL
```

```{r}
pcscores_edit_alphaPos <- pcscores_alphaPos %>% 
  filter(Indiv == "1") 

pcscores_edit_alphaPos
```

```{r}
pc_tree_d <- keep.tip(tree, tip=as.vector(pcscores_edit_d$Tip))
```

```{r}
pc_tree_CBL <- keep.tip(tree, tip=as.vector(pcscores_edit_CBL$Tip))
```

```{r}
pc_tree_alphaPos <- keep.tip(tree, tip=as.vector(pcscores_edit_alphaPos$Tip))
```

```{r}
plotTree(pc_tree_d)
```

```{r}
plotTree(pc_tree_CBL)
```

```{r}
plotTree(pc_tree_alphaPos)
```

```{r}
length(pc_tree_d$tip.label)
nrow(pcscores_edit_d)
```

```{r}
length(pc_tree_CBL$tip.label)
nrow(pcscores_edit_CBL)
```

```{r}
length(pc_tree_alphaPos$tip.label)
nrow(pcscores_edit_alphaPos)
```

```{r}
pcscores_edit_d <-
  pcscores_edit_d %>%
  remove_rownames() %>%
  column_to_rownames("FullName")
```

```{r}
pcscores_edit_CBL <-
  pcscores_edit_CBL %>%
  remove_rownames() %>%
  column_to_rownames("FullName")
```

```{r}
pcscores_edit_alphaPos <-
  pcscores_edit_alphaPos %>%
  remove_rownames() %>%
  column_to_rownames("FullName")
```

```{r}
name.check(pc_tree_d, pcscores_edit_d)
```

```{r}
name.check(pc_tree_CBL, pcscores_edit_CBL)
```

```{r}
name.check(pc_tree_alphaPos, pcscores_edit_alphaPos)
```

```{r}
pgls_Dim1_Hab_d<- gls(Dim.1 ~ Habitat, correlation = corBrownian(phy = pc_tree_d),
                      data = pcscores_edit_d, method = "ML")
summary(pgls_Dim1_Hab_d)
anova(pgls_Dim1_Hab_d)
```

```{r}
pgls_Dim1_Hab_CBL<- gls(Dim.1 ~ Habitat, correlation = corBrownian(phy = pc_tree_CBL),
                      data = pcscores_edit_CBL, method = "ML")
summary(pgls_Dim1_Hab_CBL)
anova(pgls_Dim1_Hab_CBL)
```

```{r}
pgls_Dim1_Hab_alphaPos<- gls(Dim.1 ~ Habitat, correlation = corBrownian(phy = pc_tree_alphaPos),
                      data = pcscores_edit_alphaPos, method = "ML")
summary(pgls_Dim1_Hab_alphaPos)
anova(pgls_Dim1_Hab_alphaPos)
```
