---
title: "PGLS with Full Data Set"
output: html_notebook
---


#Load in packages 
```{r}
library(rfishbase)
library (ggplot2)
library(Hmisc)
library(phytools)
library(here)
library(ape)
library(geiger)
library(nlme)
library(patchwork)
library(emmeans)
library(multcompView)
library(readxl)
library(plotly) 
library(ggthemes) 
library(naniar)
library(dplyr)
library(tidyverse)
```

#Load in csv
```{r}
fullvertmeas <- read.csv("MasterVert_Measurements.csv")
fullvertmeas
```

#Load in phylogenetic information
```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```

```{r}
tree <- read.tree(here('..', treefile))
```

```{r}
plotTree(tree)
```

```{r}
allspecies <- as.data.frame(tree$tip.label, stringsAsFactors = FALSE) 
colnames(allspecies) <- c('FullName')

allspecies
```

```{r}
allspecies <- 
  allspecies %>% separate(FullName, sep='_', into=c('Family', 'Genus', 'Species'), 
                          extra='drop', remove=FALSE)

allspecies
```

```{r}
allspecies$Tip <- seq_len(nrow(allspecies))
```

```{r}
fullvertmeas.mean <- fullvertmeas %>%
  group_by(Species, Body.Shape, Habitat_Friedman, Habitat) %>%
  summarize_at(vars(d, CBL, D_Pos, D_Ant, alpha_Pos, alpha_Ant, d_BW, D_Pos_BW, D_Ant_BW), 
               list(mean = ~mean(.x, na.rm = TRUE),
                    max = ~max(.x, na.rm = TRUE)))%>%
  ungroup() %>%
  separate(Species, sep='_', into=c("Genus", "Species"))
               
fullvertmeas.mean
```

```{r}
fullvertmeas.species <- left_join(fullvertmeas.mean, allspecies, by=c("Genus", "Species"))
fullvertmeas.species
```

```{r}
fullvertmeas.bygenus <- left_join(filter(fullvertmeas.species, is.na(Tip)),
          distinct(allspecies, Genus, .keep_all=TRUE), by=c("Genus")) %>%
  transmute(Genus=Genus, Species=Species.x, Family=Family.y, MatchSpecies=Species.y, Tip=Tip.y) 

fullvertmeas.bygenus
```

```{r}
left_join(fullvertmeas.species, fullvertmeas.bygenus,
          by = c("Genus","Species","Family")) %>%
  mutate(Tip = coalesce(Tip.x, Tip.y),
         MatchSpecies = coalesce(MatchSpecies, Species)) %>%
  select(-Tip.x, -Tip.y) %>%
  filter(!is.na(Tip)) ->
  fullvertmeas.speciesall

fullvertmeas.speciesall
```

```{r}
fullvertmeas.speciesall <- column_to_rownames(fullvertmeas.speciesall, var="FullName")

fullvertmeas.speciesall
```

```{r}
verttree <- keep.tip(tree, tip=as.vector(fullvertmeas.speciesall$Tip))
```

```{r}
plotTree(verttree)
```

#Check if tree and data match
```{r}
length(verttree$tip.label)
nrow(fullvertmeas.speciesall)
```

```{r}
name.check(verttree, fullvertmeas.speciesall)
```

#PGLS Analysis for BL mean

```{r}
pgls_dmean.BL<- gls(d_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_dmean.BL)
anova(pgls_dmean.BL)
```

```{r}
pairs(emmeans(pgls_dmean.BL, "Habitat"))
```
ANOVA is significant, but the only pair that is significant is benthic-pelagic.

```{r}
pgls_CBLmean.BL<- gls(CBL_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_CBLmean.BL)
anova(pgls_CBLmean.BL)
```

```{r}
pgls_DPosmean.BL<- gls(D_Pos_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DPosmean.BL)
anova(pgls_DPosmean.BL)
```

```{r}
pgls_DAntmean.BL<- gls(D_Ant_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DAntmean.BL)
anova(pgls_DAntmean.BL)
```

```{r}
pgls_alphaPosmean.BL<- gls(alpha_Pos_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_alphaPosmean.BL)
anova(pgls_alphaPosmean.BL)
```
The posterior cone angle is incredibly significant. 

```{r}
pairs(emmeans(pgls_alphaPosmean.BL, "Habitat"))
```
Although also here only the benthic-pelagic pair is significant.

```{r}
pgls_alphaAntmean.BL<- gls(alpha_Ant_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_alphaAntmean.BL)
anova(pgls_alphaAntmean.BL)
```
Also the anterior cone angle is very significant.

```{r}
pairs(emmeans(pgls_alphaAntmean.BL, "Habitat"))
```
Although also here only the benthic-pelagic pair is significant.

#PGLS Analysis for BL max 

```{r}
pgls_dmax.BL<- gls(d_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_dmax.BL)
anova(pgls_dmax.BL)
```
Maximum notochordal foramen diameter is significant.

```{r}
pairs(emmeans(pgls_dmax.BL, "Habitat"))
```
Super interesting here...demersal-pelagic is significant, and benthic-pelagic is ALMOST but not exactly. 

```{r}
pgls_CBLmax.BL<- gls(CBL_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_CBLmax.BL)
anova(pgls_CBLmax.BL)
```

```{r}
pgls_DPosmax.BL<- gls(D_Pos_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DPosmax.BL)
anova(pgls_DPosmax.BL)
```

```{r}
pgls_DAntmax.BL<- gls(D_Ant_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DAntmax.BL)
anova(pgls_DAntmax.BL)
```

```{r}
pgls_alphaPosmax.BL<- gls(alpha_Pos_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_alphaPosmax.BL)
anova(pgls_alphaPosmax.BL)
```
This ANOVA is very significant!

```{r}
pairs(emmeans(pgls_alphaPosmax.BL, "Habitat"))
```
Here, benthic is fully significant.

```{r}
pgls_alphaAntmax.BL<- gls(alpha_Ant_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_alphaAntmax.BL)
anova(pgls_alphaAntmax.BL)
```
This ANOVA is also very significant.

```{r}
pairs(emmeans(pgls_alphaAntmax.BL, "Habitat"))
```
Here, pelagis is fully significant.

#PGLS Analysis for BW mean

```{r}
pgls_dmean.BW<- gls(d_BW_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_dmean.BW)
anova(pgls_dmean.BW)
```

```{r}
pgls_DPosmean.BW <- gls(D_Pos_BW_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DPosmean.BW)
anova(pgls_DPosmean.BW)
```

```{r}
pgls_DAntmean.BW <- gls(D_Ant_BW_mean ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DAntmean.BW)
anova(pgls_DAntmean.BW)
```

#PGLS Analysis for BW max 

```{r}
pgls_dmax.BW<- gls(d_BW_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_dmax.BW)
anova(pgls_dmax.BW)
```

```{r}
pgls_DPosmax.BW <- gls(D_Pos_BW_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DPosmax.BW)
anova(pgls_DPosmax.BW)
```

```{r}
pgls_DAntmax.BW <- gls(D_Ant_BW_max ~ Habitat, correlation = corBrownian(phy = verttree),
                      data = fullvertmeas.speciesall, method = "ML")
summary(pgls_DAntmax.BW)
anova(pgls_DAntmax.BW)
```

#PGLS Analysis without Elops saurus

```{r}
fullvertmeas.speciesall.edit <- fullvertmeas.speciesall %>% 
  filter(Genus != "Elops")

fullvertmeas.speciesall.edit
```

```{r}
verttreeedit <- keep.tip(tree, tip=as.vector(fullvertmeas.speciesall.edit$Tip))
```

```{r}
plotTree(verttreeedit)
```

```{r}
length(verttreeedit$tip.label)
nrow(fullvertmeas.speciesall.edit)
```

```{r}
name.check(verttreeedit, fullvertmeas.speciesall.edit)
```

```{r}
pgls_dmean.BL.edit<- gls(d_mean ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_dmean.BL.edit)
anova(pgls_dmean.BL.edit)
```

```{r}
pairs(emmeans(pgls_dmean.BL.edit, "Habitat"))
```
Without Elops, pelagic is fully significant.

```{r}
pgls_alphaPosmean.BL.edit<- gls(alpha_Pos_mean ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_alphaPosmean.BL.edit)
anova(pgls_alphaPosmean.BL.edit)
```

```{r}
pairs(emmeans(pgls_alphaPosmean.BL.edit, "Habitat"))
```
Without Elops, pelagic is fully significant here.

```{r}
pgls_alphaAntmean.BL.edit<- gls(alpha_Ant_mean ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_alphaAntmean.BL.edit)
anova(pgls_alphaAntmean.BL.edit)
```

```{r}
pairs(emmeans(pgls_alphaAntmean.BL.edit, "Habitat"))
```
Without Elops, pelagic is fully significant here.

```{r}
pgls_dmax.BL.edit<- gls(d_max ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_dmax.BL.edit)
anova(pgls_dmax.BL.edit)
```

```{r}
pairs(emmeans(pgls_dmax.BL.edit, "Habitat"))
```
Pelagic is fully significant here.

```{r}
pgls_alphaPosmax.BL.edit<- gls(alpha_Pos_max ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_alphaPosmax.BL.edit)
anova(pgls_alphaPosmax.BL.edit)
```

```{r}
pairs(emmeans(pgls_alphaPosmax.BL.edit, "Habitat"))
```
Benthic is fully significant here.

```{r}
pgls_alphaAntmax.BL.edit<- gls(alpha_Ant_max ~ Habitat, correlation = corBrownian(phy = verttreeedit),
                      data = fullvertmeas.speciesall.edit, method = "ML")
summary(pgls_alphaAntmax.BL.edit)
anova(pgls_alphaAntmax.BL.edit)
```

```{r}
pairs(emmeans(pgls_alphaAntmax.BL.edit, "Habitat"))
```
Pelagic is fully significant here.

#PGLS Figures Without Elops Edit

```{r}
f1 <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=d_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

f2 <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=alpha_Pos_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

f3 <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=alpha_Ant_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)


f2 / f3 | f1
```

```{r}
f1m <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=d_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

f2m <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=alpha_Pos_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

f3m <- ggplot(fullvertmeas.speciesall, aes(x=Habitat, y=alpha_Ant_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)


f2m / f3m | f1m
```

#PGLS Figures With Elops Edit

```{r}
e1 <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=d_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

e2 <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=alpha_Pos_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

e3 <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=alpha_Ant_mean, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)


e2 / e3 | e1
```

```{r}
e1m <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=d_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

e2m <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=alpha_Pos_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

e3m <- ggplot(fullvertmeas.speciesall.edit, aes(x=Habitat, y=alpha_Ant_max, color=Habitat)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)


e2m / e3m | e1m
```











