---
title: "PGLS with PCA Dims"
output: html_notebook
---



```{r}
library(tidyverse)
library(FactoMineR)
library(factoextra)
library(patchwork)
library(ggExtra)
library(GGally)
library(rfishbase)
library (ggplot2)
library(Hmisc)
library(phytools)
library(here)
library(ape)
library(geiger)
library(nlme)
library(emmeans)
library(multcompView)
library(readxl)
library(plotly) 
library(ggthemes) 
library(naniar)
library(dplyr)
```

#Dr. Tytell's PCA Code 

```{r}
vertdata <- read_csv('All_Vert_Data.csv')
```
```{r}
head(vertdata)
```

```{r}
p1 <- ggplot(vertdata, aes(x=CBL, color=Habitat)) + geom_density()
p2 <- ggplot(vertdata, aes(x=d, color=Habitat)) + geom_density()
p3 <- ggplot(vertdata, aes(x=alpha_Ant, color=Habitat)) + geom_density()
p4 <- ggplot(vertdata, aes(x=alpha_Pos, color=Habitat)) + geom_density()

(p1 | p2) / (p3 | p4)
```

Select just the columns we need, and fill in missing positions.
```{r}
vertdata <-
  vertdata %>%
  select(c('Genus','Species', 'Indiv', 'Body.Shape', 'Habitat', 'Tip', 'Family', 'FullName',
           'Pos', 'CBL', 'd', 'alpha_Pos', 'alpha_Ant')) %>%
  # "complete" the Pos column by filling in values that are missing for Species
  # and individuals but present for others
  complete(Pos, nesting(Species, Indiv)) %>%
  arrange(Species, Indiv, Pos) %>%
  rename(alphaPos = alpha_Pos,
         alphaAnt = alpha_Ant)

vertdata
```

When we "completed" the missing positions, R filled in NAs for anything missing, 
including things like genus and species that should just repeat the same value for all positions
for the same individual. Fill in the missing qualitative pieces.
```{r}
vertdata <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  fill(Genus, Body.Shape, Habitat, Tip, Family, FullName, .direction='downup') %>%
  ungroup()

vertdata
```

Merge the genus and species columns, then push all of the categorical variables
we want to keep during the PCA to the left side of the data table.
```{r}
vertdata <-
  vertdata %>%
  unite("Species", c("Genus", "Species")) %>%
  relocate(Species, Indiv, Habitat, Tip, Pos, CBL:alphaAnt)

vertdata
```


Now we want to replace NAs with the mean value for that individual.
```{r}
vertdata_full <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  mutate_at(vars(CBL, d, alphaPos, alphaAnt), ~ replace(.x, is.na(.x), mean(.x, na.rm=TRUE))) %>%
  ungroup()

vertdata_full
```

Check how many NAs we replaced at each position.
```{r}
vertdata %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'alphaAnt')) %>%
  summarize_if(is.numeric, ~ sum(!is.na(.x)) / n()) %>%
  select_if(~ .x < 0.8)
```
This shows the fraction of each variable that is good (ie, not NA). We'll drop ones that are less than 80% good, which looks like all of the 20% and 30% measurements. alpha_Ant is different because some species just have no measurements of it.

Expand the columns so that each measurement of each variable is a different column.
```{r}
vertdata_wide <-
  vertdata_full %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'alphaAnt'))

vertdata_wide
```

Drop the 20% and 30% values, since those are missing in a lot of species. This runs
the PCA, keeping the qualitative columns 1 to 7 (`quali.sup=seq(1,7)`).
```{r}
vertdata.pca <- 
  vertdata_wide %>%
  select(-contains(c("20", "30"))) %>%
  select(-starts_with("alphaAnt")) %>%
  PCA(quali.sup=seq(1,7), scale.unit = TRUE)

vertdata.pca
```
The PCA graph shows the weights of the different variables along different axes. 
If the vectors are pointing in the same direction, then the variables are correlated
with one another. It appears that d and alphaPos are strongly correlated, while CBL
is relatively uncorrelated.

This shows how important each of the PC modes are. Variation along the first mode
explains 51.8% of the variance in the data; variance along the second mode explains
22.8% and so forth.
```{r}
fviz_eig(vertdata.pca, addlabels=TRUE)
```
The first three components get us 86% of the total variance, so that seems enough.

This extracts the PC coordinates and scores.
```{r}
vertdata.var <- get_pca_var(vertdata.pca)
vertdata.var
```
This shows how much of each of the measurement variables (rows) correspond to each 
PC axis (columns), all scaled to +-1. We need to convert this so that the "Dim.1",
"Dim.2", etc, are recorded in a variable, and put the different positions back
in another variable.
```{r}
vertdata.var$coord
```

What we're doing is kind of a rotate of the data set 90deg.
```{r}
pcweights <-
  vertdata.var$coord %>%
  # convert the coordinates into a data frame
  as.data.frame() %>%
  # the names of the variables are in the row names, so turn that into
  # a column
  rownames_to_column("varpos") %>%
  # then split variable and position into two separate columns
  separate(varpos, into=c("var", "Pos"), sep="_") %>%
  # take the Dim.1:Dim.5 columns and turn them into one long column with
  # the PC name in the "PC" column and the weight value in a "weight" column
  pivot_longer(Dim.1:Dim.5, names_to="PC", values_to="weight") %>%
  # then take the "var" column and turn each of the values into separate columns
  pivot_wider(names_from="var", values_from="weight") %>%
  mutate(Pos = as.numeric(Pos),
         PC = as.factor(PC)) %>%
  arrange(PC, Pos)

pcweights
```

The signs of the PCs are arbitrary, so if some are overall negative (ie, the mean is
less than zero), we flip the sign.
```{r}
pcweights <-
  pcweights %>%
  group_by(PC) %>%
  mutate_at(vars(CBL, alphaPos, d), ~ .x * sign(mean(.x))) %>%
  ungroup()
```

Now plot the PCs for each position and variable
```{r}
pcweights %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=d, color=PC)) +
  geom_line()
```

```{r}
pcweights %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=CBL, color=PC)) +
  geom_line()
```

```{r}
pcweights %>%
  filter(PC %in% c("Dim.1", "Dim.2", "Dim.3")) %>%
  ggplot(aes(x=Pos, y=alphaPos, color=PC)) +
  geom_line()
```


```{r}
vertdata.ind <- get_pca_ind(vertdata.pca)

pcscores <-
  vertdata.ind$coord %>%
  as.data.frame() %>%
  bind_cols(vertdata_wide)
```

```{r}
ggpairs(pcscores, columns = c("Dim.1", "Dim.2", "Dim.3"),
        mapping = aes(color = Habitat, shape = Habitat),
        upper = list(continuous = "blank"),
        diag = list(continuous = "densityDiag"))
```

Dana, this is your challenge! Do the pgls analysis on Dim.1 relative to habitat.

```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```

