---
title: "R-ifying Vertebral Measurements"
output: html_notebook
---


#Calculating body point measurements

For each scan measured, mark these points on the scan and store the values in the S column from Slicer for these points here. The values included right now are used as an example from the Sculpin01_01_V scan I measured.
```{r}
snout <- 20 #S point value of snout on Slicer 
neck <- 567 #S point value of neck on Slicer 
anal <- 159 #S point value of anal point on Slicer 
caudal <- 1536 #S point value of caudal point on Slicer 
```

Finding the standard length of the fish in the scan 
```{r}
standard_length <- abs(caudal-snout)
standard_length
```

Finding the head length of the fish in the scan
```{r}
head_length <- abs(neck-snout)
head_length
```

Finding 10% body length point if standard length > snout
```{r}
bodypoint_10 <- snout+(0.1*standard_length)
bodypoint_10
```

Finding 10% body length point if snout > standard length 
```{r}
#bodypoint_10 <- snout-(0.1*standard_length)
```

Finding 20% body length point if standard length > snout
```{r}
bodypoint_20 <- snout+(0.2*standard_length)
bodypoint_20
```

Finding 20% body length point if snout > standard length 
```{r}
#bodypoint_20 <- snout-(0.2*standard_length)
```

Finding 30% body length point if standard length > snout
```{r}
bodypoint_30 <- snout+(0.3*standard_length)
bodypoint_30
```

Finding 30% body length point if snout > standard length 
```{r}
#bodypoint_30 <- snout-(0.3*standard_length)
```

Finding 40% body length point if standard length > snout
```{r}
bodypoint_40 <- snout+(0.4*standard_length)
bodypoint_40
```

Finding 40% body length point if snout > standard length 
```{r}
#bodypoint_40 <- snout-(0.4*standard_length)
```

Finding 50% body length point if standard length > snout
```{r}
bodypoint_50 <- snout+(0.5*standard_length)
bodypoint_50
```

Finding 50% body length point if snout > standard length 
```{r}
#bodypoint_50 <- snout-(0.5*standard_length)
```

Finding 60% body length point if standard length > snout
```{r}
bodypoint_60 <- snout+(0.6*standard_length)
bodypoint_60
```

Finding 60% body length point if snout > standard length 
```{r}
#bodypoint_60 <- snout-(0.6*standard_length)
```

Finding 70% body length point if standard length > snout
```{r}
bodypoint_70 <- snout+(0.7*standard_length)
bodypoint_70
```

Finding 70% body length point if snout > standard length 
```{r}
#bodypoint_70 <- snout-(0.7*standard_length)
```

Finding 80% body length point if standard length > snout
```{r}
bodypoint_80 <- snout+(0.8*standard_length)
bodypoint_80
```

Finding 80% body length point if snout > standard length 
```{r}
#bodypoint_80 <- snout-(0.8*standard_length)
```

Finding 90% body length point if standard length > snout
```{r}
bodypoint_90 <- snout+(0.9*standard_length)
bodypoint_90
```

Finding 90% body length point if snout > standard length 
```{r}
#bodypoint_10 <- snout-(0.9*standard_length)
```

Create a data frame with the body points
```{r}
percentagebodyMarks <- data.frame("Body Percentage"=c(10,20,30,40,50,60,70,80,90), "Frame Number in Scan"=c(bodypoint_10,bodypoint_20,bodypoint_30,bodypoint_40,bodypoint_50,bodypoint_60,bodypoint_70,bodypoint_80,bodypoint_90))
percentagebodyMarks
```
Use these points to mark the vertebra at each of these frame numbers in the red view in Slicer.

#Recording Vertebral Measurement Points




