---
title: "Test phylogeny"
output: html_notebook
---

```{r}
library(tidyverse)
library(phytools)
library(here)
library(ape)
library(geiger)
library(nlme)
library(patchwork)
```

```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```

```{r}
tree <- read.tree(here('..', treefile))
```

This is our giant tree of all bony fishes.
```{r}
plotTree(tree)
```

Extract a data frame with the genus and species of each tip node in the tree.
```{r}
# first just get the tip labels. These look Family_Genus_Species_Code
allspecies <- as.data.frame(tree$tip.label, stringsAsFactors = FALSE) 
colnames(allspecies) <- c('FullName')

# split the label into columns on the underscores, but then merge genus and species again into
# a binomial separated by a space
allspecies <- 
  allspecies %>% separate(FullName, sep='_', into=c('Family', 'Genus', 'Species'), 
                          extra='drop', remove=FALSE)

# add in the tip number, which is just 1 to the number of species
allspecies$Tip <- seq_len(nrow(allspecies))
```

Load in Dana's measurements
```{r}
vertmeasfile <- here('..','Merged Data','VertMeasurements_Shape-Habitat.csv')
vertmeas <- read.csv(vertmeasfile, as.is=c(1))
```

```{r}
vertmeas
```

Get distinct species names and just extract the binomial
```{r}
vertmeas %>%
  group_by(Species, Body.Shape, Habitat) %>%
  summarize(d = mean(d),
            CBL = mean(CBL)) %>%
  ungroup() %>%
  separate(Species, sep='_', into=c("Genus", "Species")) ->
  vertmeas.species
```

Then join the big data set with Dana's data, matching by binomial
```{r}
vertmeas.species <- left_join(vertmeas.species, allspecies, by=c("Genus", "Species"))
vertmeas.species
```

Then match the tree by genus if we don't have the exact species in the tree
```{r}
left_join(filter(vertmeas.species, is.na(Tip)),
          distinct(allspecies, Genus, .keep_all=TRUE), by=c("Genus")) %>%
  transmute(Genus=Genus, Species=Species.x, Family=Family.y, MatchSpecies=Species.y, Tip=Tip.y) ->
  vertmeas.bygenus
```

Here are the closest matches we found in the tree
```{r}
vertmeas.bygenus
```

Join the genus matches into the main data set, keeping track of which species we measured and which one we matched in the tree
```{r}
left_join(vertmeas.species, vertmeas.bygenus,
          by = c("Genus","Species","Family")) %>%
  mutate(Tip = coalesce(Tip.x, Tip.y),
         MatchSpecies = coalesce(MatchSpecies, Species)) %>%
  select(-Tip.x, -Tip.y) %>%
  filter(!is.na(Tip)) ->
  vertmeas.species1
vertmeas.species1
```

**IMPORTANT**
To do the PGLS analysis, the rownames in our dataframe need to match the tip names in the tree. We've saved these values as FullName in the dataframe, so convert these back to rownames.
```{r}
vertmeas.species1 <- column_to_rownames(vertmeas.species1, var="FullName")
```

Now this pulls just the species that are in Dana's measurements out of the big phylogeny.
```{r}
verttree <- keep.tip(tree, tip=as.vector(vertmeas.species1$Tip))
```

And this plots it!
```{r}
plotTree(verttree)
```

#PGLS Analysis Attempt 
(using https://lukejharmon.github.io/ilhabela/instruction/2015/07/03/PGLS/ as a reference for the code and the PGLS process)

Check that we have the same number of tips in our tree and rows in our data frame.
```{r}
length(verttree$tip.label)
nrow(vertmeas.species1)
```

And use this extra function to make sure they match up appropriately.
```{r}
name.check(verttree, vertmeas.species1)
```

Analyze notochordal foramen diameter and its correlation with body shape.
```{r}
pgls_d_BodyShape<- gls(d ~ Body.Shape, correlation = corBrownian(phy = verttree),
                      data = vertmeas.species1, method = "ML")
summary(pgls_d_BodyShape)
```

And the centrum body length correlation with body shape.
```{r}
pgls_CBL_BodyShape<- gls(CBL ~ Body.Shape, correlation = corBrownian(phy = verttree),
                      data = vertmeas.species1, method = "ML")
summary(pgls_CBL_BodyShape)
```

And plot the data. The importance of the elongate body shape may be over stated, because there's only one example.
```{r}
p1 <- ggplot(vertmeas.species1, aes(x=Body.Shape, y=d)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

p2 <- ggplot(vertmeas.species1, aes(x=Body.Shape, y=CBL)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

p1 / p2
```

#Putting the code with the process in, regardless of whether it works (will fix once I figure out the problem)

```{r}
coef(pgls_d_BodyShape)
```








