---
title: "TestingHabitatwithFriedmanData"
output: html_notebook
---



```{r}
library(rfishbase)
library(tidyverse)
library (ggplot2)
library(Hmisc)
library(phytools)
library(here)
library(ape)
library(geiger)
library(nlme)
library(patchwork)
library(emmeans)
library(multcompView)
library(readxl)
library(plotly) 
library(ggthemes) 
```


#Load in csv
```{r}
vertmeasall <- read.csv("MasterVert_Measurements.csv")
vertmeasall
```


#Take Mean of Vert Measurements 
```{r}
vertmeasallmean <- vertmeasall %>%
  group_by(Species, Indiv) %>%
  select(SL, CBL, d, alpha_Pos, alpha_Ant, D_Pos, D_Ant) %>%
  summarize_all(mean, na.rm=TRUE) %>%
  ungroup()

vertmeasallmean
```


#Merge with Habitat Columns
```{r}
verthabitat <- read.csv("VertHabitatR.csv")

verthabitat
```

```{r}
vertmeasallmeanhabitat <-
  left_join(
  vertmeasallmean,
  
  verthabitat,
  
  by = "Species")

vertmeasallmeanhabitat
```


#Plots with Habitat_Friedman
```{r}
ggplot(vertmeasallmeanhabitat, aes(x = d, y = CBL, color = Habitat_Friedman)) + 
  geom_point()
```


```{r}
ggplot(vertmeasallmeanhabitat, aes(x = d, y = D_Pos, color = Habitat_Friedman)) + 
  geom_point()
```

```{r}
ggplot(vertmeasallmeanhabitat, aes(x = d, y = CBL, color = Habitat_FishBase)) + 
  geom_point()
```


```{r}
ggplot(vertmeasallmeanhabitat, aes(x = Habitat_Friedman, y = d, color=Habitat_Friedman)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position ="jitter", alpha=0.5)
```


```{r}
ggplot(vertmeasallmeanhabitat, aes(x = Habitat_Friedman, y = CBL, color=Habitat_Friedman)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_point(position ="jitter", alpha=0.5)
```

#Try PGLS with the Friedman Habitat Data

#Load in Phylogenetic Info
```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```


```{r}
tree <- read.tree(here('..', treefile))
```


```{r}
plotTree(tree)
```


```{r}
# first just get the tip labels. These look Family_Genus_Species_Code
allspecies <- as.data.frame(tree$tip.label, stringsAsFactors = FALSE) 
colnames(allspecies) <- c('FullName')

# split the label into columns on the underscores, but then merge genus and species again into
# a binomial separated by a space
allspecies <- 
  allspecies %>% separate(FullName, sep='_', into=c('Family', 'Genus', 'Species'), 
                          extra='drop', remove=FALSE)

# add in the tip number, which is just 1 to the number of species
allspecies$Tip <- seq_len(nrow(allspecies))
```


```{r}
vertmeasall
```


```{r}
vertmeasall %>%
  group_by(Species, Body.Shape, Habitat_Friedman) %>%
  summarise(d = mean(d),
            CBL = mean(CBL)) %>%
  ungroup() %>%
  separate(Species, sep='_', into=c("Genus", "Species")) ->
  vertmeas.species
```


```{r}
vertmeas.species <- left_join(vertmeas.species, allspecies, by=c("Genus", "Species"))
vertmeas.species
```


```{r}
left_join(filter(vertmeas.species, is.na(Tip)),
          distinct(allspecies, Genus, .keep_all=TRUE), by=c("Genus")) %>%
  transmute(Genus=Genus, Species=Species.x, Family=Family.y, MatchSpecies=Species.y, Tip=Tip.y) ->
  vertmeas.bygenus
```


```{r}
vertmeas.bygenus
```

```{r}
left_join(vertmeas.species, vertmeas.bygenus,
          by = c("Genus","Species","Family")) %>%
  mutate(Tip = coalesce(Tip.x, Tip.y),
         MatchSpecies = coalesce(MatchSpecies, Species)) %>%
  select(-Tip.x, -Tip.y) %>%
  filter(!is.na(Tip)) ->
  vertmeas.species1
vertmeas.species1
```


```{r}
vertmeas.species1 <- column_to_rownames(vertmeas.species1, var="FullName")
```


```{r}
verttree <- keep.tip(tree, tip=as.vector(vertmeas.species1$Tip))
```


```{r}
plotTree(verttree)
```


#Start PGLS Analysis with Friedman Habitat Data 
```{r}
length(verttree$tip.label)
nrow(vertmeas.species1)
```


```{r}
name.check(verttree, vertmeas.species1)
```


```{r}
pgls_d_BodyShape<- gls(d ~ Body.Shape, correlation = corBrownian(phy = verttree),
                      data = vertmeas.species1, method = "ML")
summary(pgls_d_BodyShape)
anova(pgls_d_BodyShape)
```


```{r}
pgls_CBL_BodyShape<- gls(CBL ~ Body.Shape, correlation = corBrownian(phy = verttree),
                      data = vertmeas.species1, method = "ML")
summary(pgls_CBL_BodyShape)
anova(pgls_CBL_BodyShape)
```


```{r}
p1 <- ggplot(vertmeas.species1, aes(x=Body.Shape, y=d, color=Body.Shape)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

p2 <- ggplot(vertmeas.species1, aes(x=Body.Shape, y=CBL, color=Body.Shape)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)

p1 / p2
```


```{r}
coef(pgls_d_BodyShape)
```


PGLS with Friedman Habitat Data
```{r}
pgls_d_Habitat <- gls(d ~ Habitat_Friedman, correlation = corBrownian(phy = verttree),
                      data = vertmeas.species1, method = "ML")
summary(pgls_d_Habitat)
```
Pelagic is ALMOST significant, but not really. Reef associated is, but there are only like 2 species with this category.

```{r}
pgls_CBL_Habitat <- gls(CBL ~ Habitat_Friedman, correlation = corBrownian(phy = verttree),
                        data = vertmeas.species1, method = "ML")
summary(pgls_CBL_Habitat)
```


```{r}
h1 <- ggplot(vertmeas.species1, aes(x=Habitat_Friedman, y=d, color=Habitat_Friedman)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) +
  stat_summary(fun.data = mean_sdl)


h2 <- ggplot(vertmeas.species1, aes(x=Habitat_Friedman, y=CBL, color=Habitat_Friedman)) +
  geom_point(alpha = 0.5, position = position_jitter(width=0.2)) + 
  stat_summary(fun.data = mean_sdl)

h1 / h2
```


#Try PCA with Body Shape

I tried to do a PCA but I honestly don't think I understand the math well enough to do it through code in R.























