---
title: "Morphospace plot"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(patchwork)
library(FactoMineR)
library(factoextra)
library(GGally)
library(scico)
```

#PCA with each individual measurement feature
```{r}
vertdata <- read_csv("All_Vert_Data.csv")
```

```{r}
head(vertdata)
```

```{r}
p1 <- ggplot(vertdata, aes(x=CBL, color=Habitat)) + geom_density()
p2 <- ggplot(vertdata, aes(x=d, color=Habitat)) + geom_density()
p3 <- ggplot(vertdata, aes(x=D_Pos, color=Habitat)) + geom_density()
p4 <- ggplot(vertdata, aes(x=alpha_Pos, color=Habitat)) + geom_density()

(p1 | p2) / (p3 | p4)
```

Set up for PCA

```{r}
vertdata <-
  vertdata %>%
  select(c('Genus','Species', 'Indiv', 'Body.Shape', 'Habitat', 'Tip', 'Family', 'FullName',
           'Pos', 'CBL', 'd', 'alpha_Pos', 'D_Pos')) %>%
  # "complete" the Pos column by filling in values that are missing for Species
  # and individuals but present for others
  complete(Pos, nesting(Species, Indiv)) %>%
  arrange(Species, Indiv, Pos) %>%
  rename(alphaPos = alpha_Pos,
         DPos = D_Pos)

head(vertdata)
```
Fill in NAs
```{r}
vertdata <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  fill(Genus, Body.Shape, Habitat, Tip, Family, FullName, .direction='downup') %>%
  ungroup()

head(vertdata)
```

```{r}
vertdata <-
  vertdata %>%
  unite("Species", c("Genus", "Species")) %>%
  relocate(Species, Indiv, Habitat, Tip, Pos, CBL:DPos)

head(vertdata)
```

```{r}
vertdata_full <-
  vertdata %>%
  group_by(Species, Indiv) %>%
  mutate_at(vars(CBL, d, alphaPos, DPos), ~ replace(.x, is.na(.x), mean(.x, na.rm=TRUE))) %>%
  ungroup()

head(vertdata_full)

#d.y is the parameter with the means
```

```{r}
vertdata %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'DPos')) %>%
  summarize_if(is.numeric, ~ sum(!is.na(.x)) / n()) %>%
  select_if(~ .x < 0.8)
```

```{r}
vertdata_wide <-
  vertdata_full %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'DPos',))
```

```{r}
vertdata %>%
  pivot_wider(names_from = Pos, values_from = c('CBL', 'alphaPos', 'd', 'DPos')) %>%
  summarize_if(is.numeric, ~ sum(!is.na(.x)) / n()) %>%
  select_if(~ .x < 0.8)
```

```{r}
vertdata_wide
```
Run PCA, including DPos
```{r}
vertdata.pca <-
  vertdata_wide %>%
  select(-contains(c("20", "30"))) %>%
  select(starts_with("CBL") | starts_with("d") | 
           starts_with("alphaPos") | starts_with("DPos")) %>%
  PCA(scale.unit = TRUE, ncp = ncol(.))
```

Build scores table with vertdata parameters also
```{r}
vertdata.ind <- get_pca_ind(vertdata.pca)

pcscores <-
  vertdata.ind$coord %>%
  as.data.frame() %>%
  bind_cols(vertdata_wide)

head(pcscores)
```

```{r}
ggpairs(pcscores, columns = c("Dim.1", "Dim.2"),
        aes(color=Habitat, alpha=0.5))
```

#Run PGLS on this new PCA

```{r}
treefile <- '12862_2017_958_MOESM2_ESM.tre'
```

```{r}
tree <- read.tree(here('..', treefile))
```

```{r}
plotTree(tree)
```

```{r}
allspecies <- as.data.frame(tree$tip.label, stringsAsFactors = FALSE) 
colnames(allspecies) <- c('FullName')

allspecies <- 
  allspecies %>% separate(FullName, sep='_', into=c('Family', 'Genus', 'Species'), 
                          extra='drop', remove=FALSE)

allspecies$Tip <- seq_len(nrow(allspecies))
```

```{r}
pcscores_edit <- pcscores %>% 
  filter(Indiv == "1") 

pcscores_edit
```

```{r}
pc_tree <- keep.tip(tree, tip=as.vector(pcscores_edit$Tip))
```

```{r}
plotTree(pc_tree)
```

```{r}
length(pc_tree$tip.label)
nrow(pcscores_edit)
```

```{r}
pcscores_edit <-
  pcscores_edit %>%
  remove_rownames() %>%
  column_to_rownames("FullName")
```

```{r}
name.check(pc_tree, pcscores_edit)
```

```{r}
pgls_Dim1_Hab<- gls(Dim.1 ~ Habitat, correlation = corBrownian(phy = pc_tree),
                      data = pcscores_edit, method = "ML")
summary(pgls_Dim1_Hab)
```

```{r}
anova(pgls_Dim1_Hab)
```

```{r}
pgls_Dim2_Hab<- gls(Dim.2 ~ Habitat, correlation = corBrownian(phy = pc_tree),
                      data = pcscores_edit, method = "ML")
summary(pgls_Dim2_Hab)
```

```{r}
anova(pgls_Dim2_Hab)
```


# Make morphospace

First build up the points we want to look at
```{r}
dimrangedata <- expand_grid(
  Dim.1 = seq(-5, 7, length.out = 3),
  Dim.2 = seq(-5, 7, length.out = 4),
  Dim.3 = 0,
  Dim.4 = 0,
  Dim.5 = 0
)
as.matrix(dimrangedata)
```

Function to run the reverse PCA - turn PCA coords back into real measurements.
```{r}
reverse_pca <- function(pca, U2) {
  ctr <- pca$call$centre
  ec <- pca$call$ecart.type
  
  if (is.null(dim(U2))) {
    U2 <- matrix(U2, nrow=1)
  }
  ncp1 <- ncol(U2)
  if (ncp1 > pca$call$ncp) {
    simpleError("More dimensions in new coordinates than in PCA")
  }
  S <- sqrt(pca$eig[,1])
  
  # U2 <- pca$ind$coord
  VS <- pca$var$coord
  
  V <- t(t(as.matrix(VS)) / S)
  
  X2c <- U2 %*% t(V[,1:ncp1])
  t(t(X2c) * ec + ctr)
}
```

Run the reverse PCA
```{r}
vertebrarangedata <-
  reverse_pca(vertdata.pca, as.matrix(dimrangedata)) %>%
  as_tibble() %>%
  bind_cols(dimrangedata) %>%
  pivot_longer(cols = CBL_40:alphaPos_90,
               names_to = c(".value", "pos"),
               names_sep = "_") %>%
  mutate(pos = as.numeric(pos))

head(vertebrarangedata)
```


Function to draw tho outline of a vertebra
```{r}
makeVertebra <- function(xctr, CBL, alphaPos, alphaAnt, d, w, scale=1) {
  CBL2 <- CBL/2
  d2 <- d/2
  w2 <- w/2
  ap2 <- alphaPos/2
  aa2 <- alphaAnt/2
  
  lpp <- (w - d)/2 / tan(ap2)
  lpa <- (w - d)/2 / tan(aa2)
  
  x0 <- c(-CBL2, CBL2, CBL2-lpp, CBL2-lpp, CBL2, -CBL2, -(CBL2-lpa), -(CBL2-lpa))
  y0 <- c(w2, w2, d2, -d2, -w2, -w2, -d2, d2)
  
  xf <- c(CBL2-lpp, CBL2-lpp, -(CBL2-lpa), -(CBL2-lpa))
  xf <- c(xf, xf)
  yf <- c(d2, -d2, -d2, d2)
  yf <- c(yf, yf)
  
  tibble(xout = x0*scale + xctr, 
       yout = y0*scale,
       xforam = xf*scale + xctr,
       yforam = yf*scale)
}
```

Build up vertebrae
```{r}
CBLscale <- 2

vertebrarangedata %>%
  group_by(Dim.1, Dim.2) %>%
  mutate(xctr = cumsum(1.2*CBLscale*CBL)) %>%
  mutate(xctr = xctr - mean(xctr)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(alphaPosR = alphaPos * pi/180,
         lp = (DPos - d)/2 / tan(alphaPosR/2)) %>%
  mutate(alphaAntR = if_else(lp > CBL/2,
                            2*atan((DPos-d)/2 / (CBL - lp)), alphaPosR)) %>%
  mutate(vert = pmap(list(xctr, CBLscale*CBL, alphaPosR, alphaAntR, d, DPos), makeVertebra)) %>%
  unnest(vert) ->
  vertebraoutline

head(vertebraoutline)
```
Show the vertebrae. Facets are Dim.1 and Dim.2. Color indicates foramen diameter.
```{r}
vertebraoutline %>%
  # filter((Dim.1 != 2.5) & (Dim.2 != 6)) %>%
  ggplot(aes(fill = d, group=factor(pos))) +
  geom_polygon(aes(x = xout, y = yout), color='black', fill=NA) +
  geom_polygon(aes(x = xforam, y = yforam, color = d)) +
  coord_fixed(ratio=5) +
  scale_fill_scico(palette="imola") +
  scale_color_scico(palette="imola") +
  facet_grid(Dim.2 ~ Dim.1, as.table=FALSE) +
  scale_x_continuous(breaks=NULL) +
  scale_y_continuous(breaks=NULL) +
  theme_minimal()
```

