---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(ggbeeswarm)
library(phytools)
library(patchwork)
library(here)
library(nlme)
library(ape)
library(geiger)
library(ggtree)
library(emmeans)
library(car)
library(Hmisc)
```

```{r}
citation()
print(getRversion())
```

```{r}
citation("nlme")
packageVersion("nlme")
```
```{r}
citation("ape")
packageVersion("ape")
```

```{r}
citation("geiger")
packageVersion("geiger")
```

# Load data

```{r}
vertdata_sum <- read_csv(here("output/vertdata_summary.csv")) |> 
  mutate(MatchSpecies = str_c(MatchGenus, MatchSpecies, sep = '_')) |> 
  select(-MatchGenus)
```

## Phylogeny

This is the tree from the Fish Tree of Life: https://fishtreeoflife.org/
```{r}
tree <- read.tree(here('data/actinopt_12k_treePL.tre.xz'))
```


Get the names of species from the tree.
```{r}
allspecies <- tibble(tree$tip.label)
colnames(allspecies) <- c('Species')
head(allspecies)
```

Set up the tip number (just the row)
```{r}
allspecies$Tip <- seq_len(nrow(allspecies))
```

The double species name for Fundulus confuses things, so we'll just correct it manually.
```{r}
vertdata_sum <-
  vertdata_sum |> 
  mutate(MatchSpecies = if_else(Species == 'Fundulus_heteroclitus',
                                'Fundulus_heteroclitus_heteroclitus',
                                MatchSpecies))
```

```{r}
vertdata <- left_join(vertdata_sum, allspecies, 
                                  by=c("MatchSpecies"="Species")) %>%
  select(-(ends_with(".x") | ends_with(".y")))
vertdata
```

Drop species without a match
```{r}
vertdata %>%
  filter(is.na(Tip)) %>%
  distinct(Species, .keep_all=TRUE) %>%
  select(Species, MatchSpecies, Tip, Habitat)
```

```{r}
vertdata <-
  vertdata %>%
  filter(!is.na(Tip))
```


```{r}
ourspecies <-
  vertdata %>%
  distinct(Species, .keep_all=TRUE)
```


```{r}
verttree <- keep.tip(tree, tip=as.vector(ourspecies$Tip))
```

Replace the names of species in the tree with our matched ones.

```{r}
verttree$tip.label <-
  ourspecies |> 
  arrange(Tip) |> 
  pull(Species)
```


```{r}
plotTree(verttree)
```


```{r}
vertdata_sp <- 
  vertdata %>%
  distinct(Species, .keep_all = TRUE) %>%
  mutate(rowname = Species) %>%
  column_to_rownames(var = "rowname")
```

```{r}
left_join(as_tibble(verttree),
          vertdata_sp %>%
            rownames_to_column("label") %>%
            select(label, Habitat)) %>%
  tidytree::as.treedata() %>%
  ggtree(layout = 'circular') + # geom_tiplab() +
  geom_tippoint(aes(color = Habitat))
```
Check if tree and data match
```{r}
length(verttree$tip.label)
nrow(vertdata_sp)
```

```{r}
name.check(verttree, vertdata_sp)
```

Merge the measurements and the tree
```{r}
verttree_data <-
  as_tibble(verttree) %>%
  left_join(vertdata_sp %>%
              rownames_to_column("label") %>%
              select(label, Habitat, alltaxon, Order, Family, fineness,
                     ends_with("80"), ends_with("max"), ends_with("med"), ends_with("slope"), ends_with("quad")))
```


Save out the tree data. There is something subtly different about saving the data out as a csv file and saving it as an RDS file. The base class of the tree is tbl_tree, but when we load it back in from a csv, despite having exactly the same data, it won't work with the tidytree functions. So we save it in an RDS file, which preserves the class.
```{r}
write_csv(vertdata_sp, here('output/vertdata_summary_species.csv'))
saveRDS(verttree, here('output/vert_tree.rds'))
```
