---
title: "Process FCSV files from Cymatogaster"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(here)
```

# Process Cymatogaster data

Dana processed the Cymatogaster aggregata scans separately, so we need to add them to the MasterVert_Measuremenst file.

They're in one file per vertebra:
```{r}
datadir <- 'data/Cymatogaster'
vertfiles <- c('Cymatogaster_aggregata_30.fcsv',
           'Cymatogaster_aggregata_40.fcsv',
           'Cymatogaster_aggregata_50.fcsv',
           'Cymatogaster_aggregata_60.fcsv',
           'Cymatogaster_aggregata_70.fcsv',
           'Cymatogaster_aggregata_80.fcsv',
           'Cymatogaster_aggregata_90.fcsv')
bodyfile <- 'Cymatogaster_aggregata_bodyMarks.fcsv'
```

```{r}
filenames <- map_chr(vertfiles, ~here(file.path(datadir, .x)))
```

Load in the coordinates
```{r}
coords <-
  map_dfr(filenames, ~read_csv(.x, skip=2)) %>%
  select(x,z,label)
```

```{r}
head(coords)
```
The x coordinate from the file is what we call y in the data set, and the z coordinate is x, so rename the columns. Plus extract the point number and position from the labels.
```{r}
coordsxy <-
  coords %>%
  mutate(Pos = str_extract(label, "(?<=_)\\d+"),
         PtNum = str_extract(label, "(?<=-)\\d+"),
         PtNum = as.numeric(PtNum) - 1) %>%
  rename(y = x, x = z)

coordsxy
```
Check the coordinates and make sure they make sense based on the previous data.
```{r}
coordsxy %>%
  group_by(Pos) %>%
  ggplot(aes(x = x, y = y, color=Pos, label=PtNum)) +
  geom_point() + geom_text(hjust=-0.1, vjust=0.5)
```

Make the wide data set like we have in the MasterVert_Measurements file.
```{r}
coordswide <-
  coordsxy %>%
  pivot_wider(values_from = c(x, y), id_cols = Pos,
              names_from = "PtNum", names_glue = "Pt{PtNum}{.value}")

coordswide
```

Functions for 2d and 3d distance.
```{r}
dist2 <- function(x1,y1, x2,y2) {
  sqrt((x2 - x1)^2 + (y2 - y1)^2)
}

dist3 <- function(x1,y1,z1, x2,y2,z2) {
  sqrt((x2 - x1)^2 + (y2 - y1)^2 + (z2 - z1)^2)
}
```

This calculates the different variables based on the points. (Pulled from Excel functions)
```{r}
vertmeasurements <- function(df) {
  df %>%
    mutate(CBL_raw = dist2(Pt1x, Pt1y, Pt4x, Pt4y),
           CBL2_raw = (abs(Pt1x - Pt3x) + abs(Pt2x - Pt4x))/2,
         alpha_Pos_raw = acos((dist2(Pt5x,Pt5y, Pt2x,Pt2y)^2 + dist2(Pt5x,Pt5y, Pt1x,Pt1y)^2 -
                                dist2(Pt1x,Pt1y, Pt2x,Pt2y)^2) / 
                               (2 * dist2(Pt5x,Pt5y, Pt2x,Pt2y) * dist2(Pt5x,Pt5y, Pt1x,Pt1y))) * 180/pi,
         alpha_Ant_raw = acos((dist2(Pt5x,Pt5y, Pt4x,Pt4y)^2 + dist2(Pt5x,Pt5y, Pt3x,Pt3y)^2 -
                                dist2(Pt3x,Pt3y, Pt4x,Pt4y)^2) / 
                               (2 * dist2(Pt5x,Pt5y, Pt4x,Pt4y) * dist2(Pt5x,Pt5y, Pt3x,Pt3y))) * 180/pi,
         d_raw = dist2(Pt6x,Pt6y, Pt7x,Pt7y),
         D_Pos_raw = dist2(Pt1x,Pt1y, Pt2x,Pt2y),
         D_Ant_raw = dist2(Pt3x,Pt3y, Pt4x,Pt4y))
}

reversemeasurements <- function(df) {
  df %>%
    mutate(Pt1x = D_Pos_raw / 2 / tan(alpha_Pos_raw/2 * pi/180),
           Pt2x = D_Pos_raw / 2 / tan(alpha_Pos_raw/2 * pi/180),
           Pt3x = -D_Ant_raw / 2 / tan(alpha_Ant_raw/2 * pi/180),
           Pt4x = -D_Ant_raw / 2 / tan(alpha_Ant_raw/2 * pi/180),
           Pt5x = 0,
           Pt6x = 0,
           Pt7x = 0,
           Pt1y = D_Pos_raw / 2,
           Pt2y = -D_Pos_raw / 2,
           Pt3y = D_Ant_raw / 2,
           Pt4y = -D_Ant_raw / 2,
           Pt5y = 0,
           Pt6y = d_raw / 2,
           Pt7y = -d_raw / 2)
}
```

```{r}
coordswide <-
  coordswide %>% 
  vertmeasurements()

coordswide
```
The body file contains marks for the standard length.
```{r}
bodydata <- read_csv(here(file.path(datadir, bodyfile)), skip=2) %>%
  select(x,y,z, label)
bodydata
```

These I digitized directly in the CT scan in Slicer.
```{r}
maxWidth.mm <- 175.3
maxWidthSlide <- 422
maxHeight.mm <- 492
maxHeightSlide <- 563
```

Here we pull out the standard length
```{r}
SL <-
  bodydata %>%
  mutate(Pt = str_extract(label, "(?<=Marks-)\\w+")) %>%
  select(-label) %>%
  pivot_wider(values_from = c(x, y, z), 
              names_from = "Pt", names_glue = "{Pt}{.value}") %>%
  mutate(SL = dist3(snoutx,snouty,snoutz, caudalx,caudaly,caudalz)) %>%
  pull(SL)
```

And normalize all the variables.
```{r}
coordswide <-
  coordswide %>%
  mutate(SL = SL,
         Max_BW_mm = maxWidth.mm,
         BW_slide = maxWidthSlide,
         Max_BH_mm = maxHeight.mm,
         BH_slide = maxHeightSlide,
         d = d_raw / SL,
         D_Ant = D_Ant_raw / SL,
         D_Pos = D_Pos_raw / SL,
         CBL = CBL_raw / SL,
         alpha_Pos = alpha_Pos_raw,
         alpha_Ant = alpha_Ant_raw,
         d_BW = d_raw / Max_BW_mm,
         D_Pos_BW = D_Pos_raw / Max_BW_mm,
         D_Ant_BW = D_Ant_raw / Max_BW_mm,
         `SL/Max_BW` = SL / Max_BW_mm,
         Species = "Cymatogaster_aggregata",
         Indiv = 1,
         Habitat = "pelagic",
         Pos = as.numeric(Pos))
```

# Merge with other data file

```{r}
vertmeasdata <- read_csv(here('Raw Data/MasterVert_Measurements_old.csv'))
head(vertmeasdata)
```

Check that the calculations we've done here match the data in the data table.
```{r}
checkcalc <-
  vertmeasdata %>%
  rename_with(~ str_c(.x, "_orig"), ends_with("raw")) %>%
  vertmeasurements() %>%
  mutate(alpha_Ant_raw_orig = alpha_Ant_raw_orig * 180/pi,
         alpha_Pos_raw_orig = alpha_Pos_raw_orig * 180/pi) %>%
  rename_with(~ str_c(.x, "_orig"), starts_with("Pt")) %>%
  group_by(Species, Indiv) %>%
  mutate(across(starts_with("Pt") & ends_with("x_orig"), ~ .x - Pt5x_orig),
         across(starts_with("Pt") & ends_with("y_orig"), ~ .x - Pt5y_orig)) %>%
  reversemeasurements()
```

```{r}
checkcalc %>%
  ggplot(aes(x = alpha_Ant_raw_orig, y = alpha_Ant_raw)) +
  geom_point()
```


```{r}
checkcalc %>%
  ggplot(aes(x = CBL_raw_orig, y = CBL2_raw)) +
  geom_point() +
  geom_point(aes(y = CBL_raw), color="red")
```
They match!

```{r}
checkcalc %>%
  ggplot(aes(x = Pt7y_orig, y = Pt7y)) +
  geom_point()
```

Now merge the Cymatogaster data in, and save the data file.
```{r}
vertmeasdata <-
  bind_rows(vertmeasdata, coordswide)
```

```{r}
write_csv(vertmeasdata, here('Raw Data/MasterVert_Measurements.csv'))
```

